Cooperative Kernels: GPU Multitasking for Blocking Algorithms

### 2017-07-06 write-ups/FSE/paper.tex
SQUEEZE

\myparagraph{The Needs of Irregular Data-parallel Algorithms}
%Acceleration of general-purpose computations on graphics processing
%units (GPUs) has tended to focus on \emph{regular} data-parallel
%algorithms, for which work to be processed can be evenly split between
%the workgroups that execute a GPU kernel ahead of time.  However,
Many interesting data-parallel algorithms are \emph{irregular}: the
amount of work to be processed is unknown ahead of time and may change dynamically in a workload-dependent manner.
There is growing interest in
accelerating such algorithms on
GPUs~\cite{owens-persistent,DBLP:conf/ipps/KaleemVPHP16,DBLP:conf/ipps/DavidsonBGO14,DBLP:conf/hipc/HarishN07,DBLP:journals/topc/MerrillGG15,DBLP:conf/egh/VineetHPN09,DBLP:conf/ppopp/NobariCKB12,DBLP:conf/hpcc/SolomonTT10a,DBLP:conf/popl/PrabhuRMH11,DBLP:conf/ppopp/Mendez-LojoBP12,DBLP:conf/oopsla/PaiP16,DBLP:conf/oopsla/SorensenDBGR16,DBLP:conf/egh/CedermanT08,TPO10,BNP12,Pannotia}. 
%
Irregular algorithms usually require \emph{blocking synchronization}
between workgroups.  For example, many graph algorithms employ a level-by-level
strategy, requiring a global barrier between each level, and work
stealing algorithms require each workgroup to maintain a queue, which
is typically protected by a mutex to enable stealing by other
workgroups.

@@

For example, preemption is supported by Nvidia's Pascal architecture~\cite{PascalWhitepaper}, but on a GTX Titan X (Pascal)
we still observe starvation: a global barrier
executes successfully with 56 workgroups, but deadlocks with 57
workgroups, indicating that scheduling is still unfair.

@@

\TSAdded{Cooperative kernels allow for \emph{cooperative multitasking}
  (see \mysec{\ref{sec:relatedwork}}), a multitasking approach used
  historically when preemption was not available or, as in the case of
  GPUs, too costly. Our approach avoids the cost of arbitrary
  preemption as the state of a workgroup killed via $\offerkill$ does
  not have to be saved. Previous cooperative multitasking systems have
  provided \emph{yield} semantics, where a processing unit would
  temporarily give up its hardware resource. We deviate from this
  design as, in the case of a global barrier, adopting yield would
  force the cooperative kernel to block \emph{completely} in the case
  where a single workgroup yields, stalling the kernel until the given
  workgroup resumes.  Instead, our $\offerkill$ allows a kernel to
  make progress with a smaller number of workgroups, with workgroups
  potentially joining again later via $\offerfork$.}

@@

Placing the new primitives requires manual effort, but is straightforward for the representative set of
GPU-accelerated irregular algorithms that we have ported to use cooperative kernels.  Our experiments show that the model can enable efficient multitasking of cooperative and non-cooperative tasks.

>>>>

\myparagraph{The Needs of Irregular Data-parallel Algorithms}
%Acceleration of general-purpose computations on graphics processing
%units (GPUs) has tended to focus on \emph{regular} data-parallel
%algorithms, for which work to be processed can be evenly split between
%the workgroups that execute a GPU kernel ahead of time.  However,
Many interesting data-parallel algorithms are \emph{irregular}: the
amount of work to be processed is unknown ahead of time and may change dynamically in a workload-dependent manner.
There is growing interest in
accelerating such algorithms on
GPUs~\cite{owens-persistent,DBLP:conf/ipps/KaleemVPHP16,DBLP:conf/ipps/DavidsonBGO14,DBLP:conf/hipc/HarishN07,DBLP:journals/topc/MerrillGG15,DBLP:conf/egh/VineetHPN09,DBLP:conf/ppopp/NobariCKB12,DBLP:conf/hpcc/SolomonTT10a,DBLP:conf/popl/PrabhuRMH11,DBLP:conf/ppopp/Mendez-LojoBP12,DBLP:conf/oopsla/PaiP16,DBLP:conf/oopsla/SorensenDBGR16,DBLP:conf/egh/CedermanT08,TPO10,BNP12,Pannotia}. 
%
Irregular algorithms usually require \emph{blocking synchronization}
between workgroups, e.g.\ many graph algorithms use a level-by-level
strategy, with a global barrier between levels; work
stealing algorithms require each workgroup to maintain a queue,
typically mutex-protected, to enable stealing by other
workgroups.

@@

For example, preemption is supported by Nvidia's Pascal architecture~\cite{PascalWhitepaper}, but on a GTX Titan X (Pascal)
we still observe starvation: a global barrier
executes successfully with 56 workgroups, but deadlocks with 57
workgroups, indicating unfair scheduling.

@@

\TSAdded{Cooperative kernels allow for \emph{cooperative multitasking}
  (see \mysec{\ref{sec:relatedwork}}), used
  historically when preemption was not available or too costly. Our approach avoids the cost of arbitrary
  preemption as the state of a workgroup killed via $\offerkill$ does
  not have to be saved. Previous cooperative multitasking systems have
  provided \emph{yield} semantics, where a processing unit would
  temporarily give up its hardware resource. We deviate from this
  design as, in the case of a global barrier, adopting yield would
  force the cooperative kernel to block \emph{completely} when a single workgroup yields, stalling the kernel until the given
  workgroup resumes.  Instead, our $\offerkill$ allows a kernel to
  make progress with a smaller number of workgroups, with workgroups
  potentially joining again later via $\offerfork$.}

@@

Placing the primitives manually is straightforward for the representative set of
GPU-accelerated irregular algorithms we have ported so far.  Our experiments show that the model can enable efficient multitasking of cooperative and non-cooperative tasks.
